---
title: "Lab 3 Report"
author: "Alec White"
date: "`r format(Sys.Date())`"
output: github_document
---

Instructions for this lab report can be found on [Your assignment section of Lab Assignment 3 Introduction to ggplot2](https://biol275-msum.github.io/introduction-to-ggplot2.html#your-assignment) on the lab website.

## Fireflies

# load packages -----------------------------------------------------------
>A. insert Rcode chunk 
```{r}

firefly_data <- read_csv("https://whitlockschluter.zoology.ubc.ca/wp-content/data/chapter02/chap02q19FireflySpermatophoreMass.csv")

firefly_data

ggplot(data = firefly_data)

ggplot(data = firefly_data) +
  geom_histogram(mapping = aes(x = spermatophoreMass))

ggplot(data = firefly_data) +
  geom_histogram(mapping = aes(x = spermatophoreMass),
                 boundary = 0, closed = "left", 
                 fill = "#C5351B", color = "black") +
  labs(x = "Spermatophore mass", y = "Frequency (number of eggs)") +
  scale_y_continuous(breaks = seq(0, 10, 1), limits = c(0, 10), 
                     expand = expansion(mult = 0)) +
  scale_x_continuous(breaks = seq(0, .20, .05)) +
  theme_classic() +
  theme(
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black", size = rel(1))
  )

```
 
\_\_\_

> B. For part (A) I used A histogram due to the nature of the data being numerical, which uses only one variable but has a large range of data.

\_\_\_

> C. The frequency distributon shows to be a bell curve while ignoring the one outlier.

\_\_\_

> D. The largest measurement in the frequency distribution is considered the Mode. 


\_\_\_

## Bird orders

I read the data using the code provided by the instructor. The dataset is from the auk package and was originally named ebird_taxonomy. I converted it to a tibble, removed non-species taxa, and named the new dataset birds.

```{r}
library(auk)                          # load the auk package
birds <- ebird_taxonomy %>%           # start with the ebird_taxonomy data
  as_tibble() %>%                     # tibbles print better in the console
  filter(category == "species")       # remove non-species taxa
```

> E. There are 10,721 bird species, which was found using the count(species) command on the birds dataset.

\_\_\_

> H. There are 41 orders of birds, which was found using the count(orders) command on the birds dataset.

\_\_\_

> F. Insert an R code chunk and create a graph depicting the distribution of orders in the birds dataset. Sort the orders with the most frequent on the left. It should have legible text and appropriate axis labels.
```{r}

library(auk)                          # load the auk package
birds <- ebird_taxonomy %>%           # start with the ebird_taxonomy data
  as_tibble() %>%                     # tibbles print better in the console
  filter(category == "species")       # remove non-species taxa

birds

distinct(birds, order)

count(birds, order)

ggplot(data = birds)

ggplot(data = birds) +
  geom_bar(mapping = aes(x = order))

ggplot(data = birds) +
  geom_bar(mapping = aes(x = fct_infreq(order)))

ggplot(data = birds) +
  geom_bar(mapping = aes(x = fct_infreq(order)), fill = "#C5351B")

ggplot(data = birds) +
  geom_bar(mapping = aes(x = fct_infreq(order)), fill = "#C5351B", 
           width = .8) +
  labs(x = "Order", y = "Frequency (log number of bird species)", 
       title = "Number of bird species in each order", 
       subtitle = "*numbers are in log scale to better show data.") +
  scale_y_log10() +
  theme_classic(base_size = 12) +
  theme(
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black", size = rel(1)),
    axis.text.x = element_text(angle = 60, hjust = 1),
    axis.ticks.x = element_blank()
  )
```

\_\_\_

> G. For the birds data-set I created a bar graph due to its data being of categorical nature. The different bars are much more efficient at expressing all of the different orders being examined in this data.

\_\_\_

## Links to peer review

I reviewed Lukas Robinson's lab report at this URL: 

Lukas Robinson reviewed my lab report at this URL: https://github.com/BIOL275-MSUM/lab-3-intro-to-ggplot-AWhite3385/issues/1

## Session Info

Here is my session information:

```{r}
sessioninfo::session_info()
```
